---
title: "p8105_hw2_zg2627"
author: "Zimu Guan"
date: "2026-10-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r, message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(janitor)

# Import the NYC subway entrance/exit data
subway_data =
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  clean_names()
```
```{r}
# Keep the variables needed for the analysis and recode entry as logical
subway_clean =
  subway_data |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entrance_type,
    entry,
    vending,
    ada
  ) |>
  mutate(
    entry = entry == "YES"
  )

dim(subway_clean)
```

The cleaned subway dataset contains information on subway entrances and exits, including subway line and station name, station coordinates, routes served, entrance type, whether entry is allowed, vending availability, and ADA accessibility. During cleaning, I retained the variables relevant to the analysis and converted entry from a character variable ("YES" / "NO") to a logical variable. The resulting dataset contains 1868 observations and 19 variables. The data are not fully tidy because route information is stored across multiple columns (route1 through route11).

```{r}
# Count distinct subway stations
subway_clean |>
  distinct(line, station_name) |>
  nrow()

# Count distinct ADA-compliant stations
subway_clean |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()

# Calculate the proportion of entrances without vending that allow entry
subway_clean |>
  filter(vending == "NO") |>
  summarize(
    proportion_entry = mean(entry)
  )
```

There are 465 distinct stations, of which 84 are ADA compliant. Among entrances and exits without vending, approximately 37.7% allow entry.

```{r}
# Reshape route variables into a tidy format
subway_tidy =
  subway_clean |>
  mutate(
    across(route1:route11, as.character)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  filter(!is.na(route_name))
```


```{r}
# Keep stations served by the A train
a_train_stations =
  subway_tidy |>
  filter(route_name == "A")

# Count distinct stations served by the A train
a_train_stations |>
  distinct(line, station_name) |>
  nrow()

# Count ADA-compliant stations served by the A train
a_train_stations |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

To address this issue, I reshaped route1 through route11 so that route number and route name are stored as separate variables. The A train serves 60 distinct stations, 17 of which are ADA compliant.


## Problem 2

```{r}
library(readxl)

# Import and clean the Mr. Trash Wheel data
mr_trash =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N743"
  ) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )
```

```{r}
# Import and clean the Professor Trash Wheel data
professor_trash =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M140",
    na = c("", "dive")
  ) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  )
```

```{r}
# Import and clean the Gwynnda data
gwynnda_trash =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = "A2:L395"
  ) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Gwynnda"
  )
```

```{r}
# Combine the three Trash Wheel datasets
trash_wheel_data =
  bind_rows(
    mr_trash,
    professor_trash,
    gwynnda_trash
  )
```

```{r}
# Calculate the total weight collected by Professor Trash Wheel
professor_total_weight =
  trash_wheel_data |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  summarize(
    total_weight = sum(weight_tons, na.rm = TRUE)
  ) |>
  pull(total_weight)

# Calculate cigarette butts collected by Gwynnda in June 2022
gwynnda_june_2022 =
  trash_wheel_data |>
  filter(
    trash_wheel == "Gwynnda",
    year == 2022,
    month == "June"
  ) |>
  summarize(
    total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE)
  ) |>
  pull(total_cigarette_butts)
```

The combined Trash Wheel dataset contains `r nrow(trash_wheel_data)` observations and `r ncol(trash_wheel_data)` variables from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Key variables include dumpster number, month and year of collection, trash weight and volume, and counts of different types of debris such as plastic bottles, polystyrene, cigarette butts, plastic bags, and wrappers. For the available data, Professor Trash Wheel collected a total of `r professor_total_weight` tons of trash. Gwynnda collected a total of `r format(gwynnda_june_2022, big.mark = ",", scientific = FALSE)` cigarette butts in June 2022.

## Problem 3

```{r}

```





